\bf Manifold Clustering in the Generalized Random Dot Product Graph

### 2022-05-04 draft/draft.tex
still working on it

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

@@

% Pandoc citation processing

@@

\hypertarget{latent-structure-block-models}{%
\section{Latent Structure Block
Models}\label{latent-structure-block-models}}

All block models are latent structure models.

@@

In this section, we consider the following scenario: Suppose that each
community is represented by a manifold \(\mathcal{M}_k\),
\(k \in \{1, ..., K\}\) in the latent space of a RDPG or GDRPG, and the
manifolds do not intersect each other. Define
\(\delta = \min\limits_{k \neq \ell} \min\limits_{x \in \mathcal{M}_k, y \in \mathcal{M}_\ell} \|x - y\|\),
the minimum distance between two manifolds. We assume that
\(\delta > 0\), i.e., the manifolds do not intersect.

@@

\begin{algorithm}[h]
\DontPrintSemicolon
\SetAlgoLined
\KwData{Adjacency matrix $A$, number of communities $K$, embedding dimensions $p$ and $q$.}
\KwResult{Community assignments $z_1, ..., z_n \in \{1, ..., K\}$.}
Compute $\hat{X}$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.\;
Apply single linkage clustering with $K$ communities on $\hat{X}$\;
\caption{ASE clustering for nonintersecting communities.}
\end{algorithm}

Consider the specific case in which each
\(\mathcal{M}_k \subset \mathbb{R}^d\) is a one-dimensional manifold.
Let \(F_k\) be a probability distribution with support
\(\mathcal{M}_k\). Then we define a mixture model as follows:

@@

In this setting, although each manifold is in \(d\)-dimensional space,
since the manifolds are one-dimensional, it is possible to consider an
ordering of points sampled on the manifold. In the original latent space
in which there is no noise, the condition for which community detection
is possible via \(\eta\)-neighborhood graph clustering or single linkage
clustering is satisfied if each pair of adjacent order statistics on
each manifold is less than \(\delta\) apart.

\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds without noise]
\label{nonintersect-no-noise}
Let $x_1, ..., x_n$ be points sampled on $K$ one-dimensional manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$. 
Suppose $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \| x_i - x_j \| > 0$. 
Let $A_n$ be the event such that $\max_k \max_i \|x_{(i+1)}^{(k)} - x_{(i)}^{(k)}\| < \delta$, where $x_{(i)}^{(k)}$ is the $i^{th}$ order statistic on the $k^{th}$ manifold. 
Then for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(A_n) > 1 - \epsilon$. 
\end{theorem}

Sampling a graph from this latent configuration and constructing an ASE
from the graph introduces error relative to the original latent vectors.
If the errors are small, which is guaranteed by the consistency property
of the ASE for sufficiently large \(n\), then sufficient separation
between embedding points belonging to different manifolds is still
guaranteed.

\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds with noise]
\label{nonintersect-with-noise}
Suppose the same setup as in theorem \ref{nonintersect-no-noise}. 
Define $y_i = x_i + e_i$ and $B_n$ as the event such that $\|y_{(i+1)}^{(k)} - y_{(i)}^{(k)}\| < \|y_{(i)}^{(k)} - y_{(j)}^{(\ell)}\|$ for any $k \neq \ell$, where, $y_{(i)}^{(k)}$ is the $i^{th}$ order statistic on the $k^{th}$ manifold plus its corresponding noise. 
Then if $\max_i \|e_i\| < \delta / 3$, for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(B_n) > 1 - \epsilon$. 
\end{theorem}

\begin{theorem}[Community detection for one-dimensional nonintersecting manifold block models]
\label{nonintersect-mbm}
Suppose we have the same setup as in theorem \ref{nonintersect-no-noise}. 
Let $p, q \geq 0$ such that $p + q = d$ where $d$ is the dimension of the range of each $g_k$. 
Suppose we draw $A \sim \mathrm{GRDPG}_{p,q}(X; \rho_n)$ for some sparsity parameter $\rho_n \in (0, 1]$, and let $\hat{X}$ be the ASE of $A$. 
Then single linkage clustering produces zero community detection error with high probability as $n \to \infty$. 
\end{theorem}

@@

To prove theorems \ref{nonintersect-no-noise},
\ref{nonintersect-with-noise}, and \ref{nonintersect-mbm}, we first need
to introduce some theory for order statistics. Let
\(D_i^{(k)} = X_{(i+1)}^{(k)} - X_{(i)}^{(k)}\), where \(X_{(i)}^{(k)}\)
is the \(i^{th}\) order statistic of univariate sample
\(X_1^{(k)}, ..., X_n^{(k)} \stackrel{\mathrm{iid}}{\sim}F^{(k)}\) and
\(F^{(k)}\) is a continuous distribution on manifold \(\mathcal{M}_k\).
If \(\max_i D_i^{(k)} < \delta\), then there is sufficient separation of
points between each manifold. Then it is sufficient to quantify
\(P(\max_i D_i^{(k)} > \delta)\) for each \(k\) as a function of \(n\)
and \(\delta\) and show that this converges to zero as \(n\) grows to
\(\infty\).

For ease of notation, we drop the superscript denoting
community/manifold membership until lemma 6. Denote \(f(x)\) as the
density of each \(F\), \(g_i(x)\) as the density of \(X_{(i)}\),
\(g_{ij}(x, y)\) as the joint density of \(X_{(i)}, X_{(j)}\), and
\(h_i(d)\) as the density of \(D_i\) (with corresponding capital letters
for the cumulative distribution functions).

The following are taken as given\footnote{\url{https://en.wikipedia.org/wiki/Order_statistic}}:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \(g_i(x) = \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x))^{n-i} f(x)\).
\item
  \(g_{ij}(x, y) = \frac{n!}{(i-1)! (j-i-1)! (n-j)!} (F(x))^{i-1} (F(y) - F(x))^{j-i-1} (1 - F(y))^{n-j} f(x) f(y)\).
\item
  By convolution,
  \(h_i(d) = \int_0^{1} g_{i, i+1} (x, x + d) \mathrm{d}x\).
\end{enumerate}

\begin{lemma}[The probability density function of $D_i$]
Suppose $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ and the support of $F$ is the unit interval. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then the density of each $D_i$ is:

\begin{equation}
\label{eq:pdf}
h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \mathrm{d}x 
\end{equation}
\end{lemma}

\begin{proof}
This is a direct consequence of 2 and 3. 
We also note that because the support of $X_i$ is $[0, 1]$, 
the integral only needs to be evaluated from $0$ to $1 - d$ 
because of the $f(x+d)$ and $1 - F(x+d)$ terms.
\end{proof}

\begin{lemma}[The cumulative distribution function of $D_i$]
Suppose $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ and the support of $F$ is the unit interval. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then the distribution function of each $D_i$ is:

\begin{equation}
\label{eq:cdf}
P(D_i < \delta) = H_i(\delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x
\end{equation}
\end{lemma}

\begin{proof}
$$
\begin{aligned}
H_i(\delta) & = \int_x^{x+\delta} h_i(d) \mathrm{d}d \\
& = \int_x^{x+\delta} \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} ((F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \mathrm{d}x \mathrm{d}d \\
& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_x^{x+\delta} (1 - F(x+d))^{n-i-1} f(x+d) \mathrm{d}d \mathrm{d}x \\
& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_{F(x)}^{F(x+\delta)} (1 - u)^{n-i-1} \mathrm{d}u \mathrm{d}x \\
& = \int_0^{1} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} f(x) ((1 - F(x))^{n-i} - (1 - F(x + \delta))^{n-i}) \mathrm{d}x \\ 
& = \int_0^1 g_i(x) \mathrm{d}x - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x \\
& = 1 - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x \\
\end{aligned}
$$

Because of the $x + \delta$ term and the fact that the support is $[0, 1]$, the integrand is zero above $1 - \delta$, so we are left with

$$
\begin{aligned}
& = 1 - \int_0^{1 - \delta} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x.
\end{aligned}
$$
\end{proof}

We now focus on the case where points are sampled uniformly on the unit
interval to build up to more general distributions.

\begin{lemma}[Differences between order statistics of a uniform distribution]
If $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Uniform}(0, 1)$, then each $D_i \sim \mathrm{Beta}(1, n)$.
\end{lemma}

\begin{proof}
We begin with Eq. (\ref{eq:pdf}), plugging in $f(x) = 1$ and $F(x) = x$:

$$
h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \mathrm{d}x.
$$

Then we proceed with integration by parts, setting 
$u = x^{i-1} \implies du = (i-1) x^{i-2}$ and 
$dv = (1-x-d)^{n-i-1} dx \implies v = -\frac{1}{n-i} (1-x-d)^{n-i-1}$. 
Note that $u v |_0^{1-d} = 0$ in this case. This yields

$$
= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \mathrm{d}x.
$$

Then applying integration by parts again until the $x^p$ term disappears, we get:

$$
\begin{aligned}
& = \frac{n!}{(i-1)! (n-i-1)!} \frac{(i-1)!}{(n-i) \cdots (n-2)} \int_0^{1-d} (1-x-d)^{n-2} \mathrm{d}x \\
& = -\frac{n (n-1)}{n-1} (1-x-d)^{n-1} \Big|_0^{1-d} \\
& = n (1 - d)^{n-1}.
\end{aligned}
$$

This the density function for $\mathrm{Beta}(1, n)$, completing the proof.
\end{proof}

\begin{lemma}
Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Uniform}(0, 1)$. 
Then for any $\epsilon$ and $\delta > 0$, 
there exists an $N = O \big(\frac{-\log \epsilon}{\delta} \big)$ such that 
$P(\max_i X_{(i+1)} - X_{(i)} < \delta) \geq 1 - \epsilon$ when $n > N$.
\end{lemma}

\begin{proof}
Since $X_{(i+1)} - X_{(i)} = D_i \sim \mathrm{Beta}(1, n)$, 
$P(X_{(i+1)} - X_{(i)} < \delta) = 1 - (1 - \delta)^n $. This yields

$$
\begin{aligned}
P(\max_i D_i < \delta) & \geq (P(D_i < \delta))^{n-1} \\
& = (1 - (1 - \delta)^n)^{n-1} \\
& \approx e^{-n \exp(-n \delta)}.
\end{aligned}
$$

In the limit $n \to \infty$, this goes to 1.
\end{proof}

Now we extend this lemma to general distributions on the unit interval.

\begin{lemma}
\label{lem:lim-d}
Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then for any $\epsilon > 0$, there exists $N > 0$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
\end{lemma}

\begin{proof}
We start with Eq. (\ref{eq:cdf}):

$$P(D_i \leq \delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x.$$

Making the approximation $F(x+\delta) \approx F(x) + \delta f(x)$ 
and bounding $f(x) \geq a$, we get:

$$P(D_i \leq \delta) \geq 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x) - a \delta)^{n-i} f(x) \mathrm{d}x.$$

Then making the substitution $u = F(x) \implies du = f(x) \mathrm{d}x$, we obtain

$$1 - \int_0^{F(1-\delta)} \frac{n!}{(n-i)! (i-1)!} u^{i-1} (1 - u - a \delta)^{n-i} \mathrm{d}u$$

Evaluating the integral yields

$$P(D_i < \delta) = 1 - (1 - a \delta)^n + (1 - F(1-\delta) - a \delta)^n.$$

Then as before,

$$
\begin{aligned}
P(\max_i D_i < \delta) & = P(\text{all } D_i < \delta) \\
& = 1 - P(\text{some } D_i > \delta) \\
& \geq 1 - \sum_i^{n-1} P(D_i > \delta) \\
& = 1 - (n - 1) (1 - a \delta)^n + (n - 1) (1 - F(1 - \delta) - a \delta)^n
\end{aligned}
$$

Since $(1 - a \delta) < 1$ and $(1 - F(1-\delta) - a \delta) < 1$, this converges to $1$ in the limit $n \to \infty$. 
Then for any $\epsilon \in (0, 1)$, we can always solve for an $n$ such that $1 - (n-1) (1 - a \delta)^n (n-1) (1 - F(1 - \delta) - a \delta)^n \geq 1 - \epsilon$ holds. 
\end{proof}

\begin{proof}[Proof of theorem \ref{nonintersect-no-noise}]
This theorem requires two additional conditions: 
First, each $n_k \to \infty$ as $n \to \infty$, and second, the number of manifolds $K$ remains fixed. 
These are satisfied by the model assumptions. 
Then this is a direct consequence of Lemma \ref{lem:lim-d}. 
Denoting $m = \min_k n_k$ as the sample size of the manifold with the fewest points,

$$
\begin{aligned}
P(\max_{k,i} \|X_{(i+1)}^{(k)} - X_{(i)}^{(k)}\| < \delta) & = P(\forall k, i ~ D_i^{(k)} < \delta) \\
& \geq \prod_{k=1}^K P(D_i^{(k)} < \delta) \\
& \geq (1 - (m-1) (1 - a \delta)^{m} + (m-1) (1 - F(1-\delta) - a \delta)^{m})^K
\end{aligned}
$$
Since as $n \to \infty$, $m \to \infty$ as well, and $K$ is fixed, this approaches $(1)^K = 1$. Then setting this $\geq 1 - \epsilon$, it is always possible to solve this for $m$. 
\end{proof}

\begin{proof}[Proof of theorem \ref{nonintersect-with-noise}]
Let $\max_i \|e_i\| \leq \xi$. 
\end{proof}

\begin{proof}[Proof of theorem \ref{nonintersect-mbm}]
\end{proof}

Here, we extend the results on the line to the unit hypercube. The
following lemma is based on lemma 2 of
\citet{trosset2020rehabilitating}.

\begin{lemma}
\label{thm:multidim}
Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. 
Define $E_n$ as the event that an $\eta$-neighborhood graph constructed from the sample is connected. 
Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon \eta^r / r^{r/2}}{\log (1 - \frac{a \eta^r}{r^{r / 2}})} \bigg)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
\end{lemma}

\begin{proof}[Proof (sketch)]
Divide the hypercube $[0, 1]^r$ into a grid of sub-hypercubes of side length at most $\eta / \sqrt{r}$. 
$E_n$ is satisfied if each sub-hypercube contains at least one $X_i$ from the sample.

$$
\begin{aligned}
P(E_n) & = 1 - P(\text{some cells don't contain } X_i) \\
& \geq 1 - \sum_k^{\lceil \sqrt{r} / \eta \rceil^r} \prod_i^n P(X_i \text{ is not in the } k^{th} \text{ hypercube}) \\
& \geq 1 - \lceil \sqrt{r} / \eta \rceil^r (1 - a \eta^r / r^{r/2})^n
\end{aligned}
$$

Setting this $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
\end{proof}

>>>>

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}

@@

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

@@

\hypertarget{generalized-random-dot-product-graphs-with-community-structure}{%
\section{Generalized Random Dot Product Graphs with Community
Structure}\label{generalized-random-dot-product-graphs-with-community-structure}}

All Bernoulli graphs are generalized random dot product graphs. Whether
this is useful for inference depends on the structure of the latent
space. In the case of the \(\text{Erd\"{o}s-R\'{e}nyi}\) model, SBM,
DCBM, and PABM, the latent structure is linear, and the linearity can be
exploited for community detection and parameter estimation. In this
section, we discuss general, often nonlinear latent structure models,
focusing on those with community structure.

To motivate this, consider a generalization of the
\(\text{Erd\"{o}s-R\'{e}nyi}\) model. Recall that when viewed as an
RDPG, the latent space of an \(\text{Erd\"{o}s-R\'{e}nyi}\) model
consists of one point in Euclidean space. In the following example,
instead of fixing the edge probability, it is sampled from a
distribution in such a way that when viewed as an RDPG, the latent space
consists of a curve.

\begin{example}[Hierarchical $\text{Erd\"{o}s-R\'{e}nyi}$ model]
In the $\text{Erd\"{o}s-R\'{e}nyi}$ model, the edge probability matrix has a fixed value $[P_{ij}] \equiv p \in [0, 1]$. 
Suppose instead that each $P_{ij}$ is drawn from a distribution with density $f(p) = \frac{1}{\frac{\pi}{2} - 1} \Big(\frac{1}{\sqrt{1 - p^2}} - 1 \Big)$.

Then it can be shown that this model is equivalent to a random dot product graph with a two dimensional latent space in which latent vectors are sampled uniformly along the curve $g(t) = \begin{bmatrix} \cos(\frac{\pi}{2} t) & \sin(\frac{\pi}{2} t) \end{bmatrix}^\top$, $0 \leq t \leq 1$. 
\end{example}

@@

In this section, we consider the following scenario: Suppose that each
community is represented by a closed manifold \(\mathcal{M}_k\),
\(k \in \{1, ..., K\}\) in the latent space of a RDPG or GDRPG. Define
\(\delta = \min\limits_{k \neq \ell} \min\limits_{x \in \mathcal{M}_k, y \in \mathcal{M}_\ell} \|x - y\|\),
the minimum distance between two manifolds. We assume that
\(\delta > 0\), i.e., the manifolds do not intersect.

@@

\begin{algorithm}[h]
\DontPrintSemicolon
\SetAlgoLined
\KwData{Adjacency matrix $A$, number of communities $K$, embedding dimensions $p$ and $q$.}
\KwResult{Community assignments $z_1, ..., z_n \in \{1, ..., K\}$.}
Compute $\hat{X}$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.\;
Apply single linkage clustering with $K$ communities on $\hat{X}$.\;
\caption{ASE clustering for nonintersecting communities.}
\end{algorithm}

Let \(F_k\) be a probability distribution with support
\(\mathcal{M}_k\). Then we define a mixture model as follows:

@@

\begin{theorem}[Community detection for nonintersecting manifolds without noise]
\label{nonintersect-no-noise}
Let $x_1, ..., x_n$ be points sampled on $K$ manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$. 
Suppose $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \| x_i - x_j \| > 0$. 
Define $A_n$ as the event that for a $\eta$-neighborhood graph constructed from the sample $x_1, ..., x_n$ consists of exactly $K$ disconnected subgraphs that map on exactly to each manifold for some $\eta \in (0, \delta)$. 
Then for any $\epsilon \in (0, 1)$, there exists an $N$ such that when $n > N$, $A_n$ occurs with high probability. 
\label{nonintersect-no-noise}
\end{theorem}

@@

In order to prove theorem \ref{nonintersect-no-noise}, we first
establish that the density of points within one manifold is sufficiently
dense. The following lemma is based on lemma 2 of
\citet{trosset2020rehabilitating}.

\begin{lemma}
\label{lem:one-hypercube}
Let $x_1, ..., x_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. 
Define $H_n$ as the event that an $\eta$-neighborhood graph constructed from the sample is connected for any $\eta > 0$. 
Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon + r \log \eta - \frac{r}{2} \log r}{\log (1 - a \eta^r r^{-r / 2})} \bigg)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
\end{lemma}

\begin{proof}
Divide the hypercube $[0, 1]^r$ into a grid of sub-hypercubes of side length at most $\eta / \sqrt{r}$. 
$E_n$ is satisfied if each sub-hypercube contains at least one $X_i$ from the sample.

$$
\begin{aligned}
P(H_n) & = 1 - P(\text{some cells don't contain } X_i) \\
& \geq 1 - \sum_m^{\lceil \sqrt{r} / \eta \rceil^r} \prod_i^n P(X_i \text{ is not in the } k^{th} \text{ hypercube}) \\
& \geq 1 - \lceil \sqrt{r} / \eta \rceil^r (1 - a \eta^r / r^{r/2})^n,
\end{aligned}
$$
which approaches $1$ as $n \to \infty$. 
Setting this quantity as  $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
\end{proof}

\begin{lemma}
\label{lem:K-hypercubes}
Let there be $K \geq 2$ hypercubes $\mathcal{C}_1, ..., \mathcal{C}_K$ of dimension $r_1, ..., r_K$ in $\mathbb{R}^d$ such that $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{C}_k, x_j \in \mathcal{C}_\ell} \|x_i - x_j\| > 0$. 
Let $F_k$ be a distribution with support $C_k$ such that its density $f_k$ is nonzero on the support. Let $a = \min_k \min_t f_k(t) > 0$. 
Define a mixture model as follows:

\begin{enumerate}
  \item Draw labels $z_1, ..., z_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\alpha_1, ..., \alpha_K)$. 
  \item Draw latent vectors each as $x_i \stackrel{\mathrm{ind}}{\sim}F_{z_i}$.
\end{enumerate}

Define $H'_n$ as the event that an $\eta$-neighborhood graph constructed from this sample consists of exactly $K$ disconnected subgraphs for any $\eta \in (0, \delta)$. 
Then for any $\epsilon \in (0, 1)$, there exists an $N = O \Big( \frac{\log(1 - (1 - \epsilon)^{1/K}) + d \log \eta - \frac{d}{2} \log d}{\alpha_{min} \log(1 - a \eta^d d^{-d/2})} \Big)$ such that when $n > N$, $P(E_n) > 1 - \epsilon$. 
\end{lemma}

\begin{proof}
Let $H^{(k)}$ be the event that lemma \ref{lem:one-hypercube} holds for $\mathcal{C}_k$. Then $H^{(k)} = H_{n_k}$ where $H_n$ is defined as in \ref{lem:one-hypercube}. 
Then $P(H'_n) = P(H_{n_1} \text{ and } ... \text{ and } H_{n_K})$.

$$
\begin{aligned}
P(H'_n) & = \prod_k P(H_{n_k}) \\
& \geq \prod_k 1 - \lceil \sqrt{r_k} / \eta \rceil^{r_k} (1 - a \eta^{r_k} r_k^{-r_k/2})^{n_k} \\
& \geq \prod_k 1 - \lceil \sqrt{d} / \eta \rceil^{d} (1 - a \eta^{d} d^{-d/2})^{\alpha_{\min} n} \\
& \geq \big(1 - \lceil \sqrt{d} / \eta \rceil^{d} (1 - a \eta^{d} d^{-d/2})^{\alpha_{\min} n} \big)^K, \\
\end{aligned}
$$
which approaches 1 as $n \to \infty$. 
Setting this quantity to $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
\end{proof}

\begin{proof}[Proof of theorem \ref{nonintersect-no-noise}]
\end{proof}
